Hydrodynamics of a Semipermeable Inextensible Membrane Under Flow and Confinement

### 2021-03-18 vesicle_code/docs/permeableVes/main.tex

% At least three keywords are required at submission. Please provide three to five keywords, separated by the pipe symbol.
\keywords{Semipermeable membrane$|$ Vesicle $|$ Extreme confinement $|$
Boundary integral equations $|$ Stokes flow}

@@

\setboolean{displaywatermark}{false}

@@

Cells exhibit a dynamic volume-to-surface ratio in 
many physiological processes, such as 
cell migration in confined microenvironments~\cite{StrokaJiangChenEtAl2014_Cell, papadopoulos2008aquaporins}, 
dialysis~\cite{wan2006}, cell desiccation~\cite{vogl2014effect}, cell division~\cite{OdermattMiettinenKangEtAl2020_bioRxiv} and
apoptosis~\cite{marchetti1996mitochondrial}.
%osmotic flow across the membrane.
While osmotic stress  is more effective and efficient than mechanical stress in contributing to change in cellular volume, our results suggest that the water permeation due to the capillary pressure may be non-negligible over long times or under strong confinement in simple configurations. 
To examine if this is the case in the physiological processes,  currently we are examining the combined effects of osmotic and mechanical stresses on water permeation in the small Peclet number limit, which is relevant to cellular migration in the presence of both confinement and a solute gradient~\cite{StrokaJiangChenEtAl2014_Cell} and has been studied numerically by Jaeger {\em et al.}~\cite{jae-car-med-try1999}

>>>>

% At least three keywords are required at submission. Please provide three to five keywords, separated by the pipe symbol.
\keywords{Semipermeable membrane $|$ Vesicle $|$ Extreme confinement $|$
Boundary integral equations $|$ Stokes flow}

@@

\setboolean{displaywatermark}{true}

@@

Cells exhibit a dynamic volume-to-surface ratio in 
many physiological processes, such as 
cell migration in confined microenvironments~\cite{StrokaJiangChenEtAl2014_Cell, papadopoulos2008aquaporins}, 
dialysis~\cite{wan2006}, cell desiccation~\cite{vogl2014effect}, cell division~\cite{OdermattMiettinenKangEtAl2020_bioRxiv}, and
apoptosis~\cite{marchetti1996mitochondrial}.
%osmotic flow across the membrane.
While osmotic stress  is more effective and efficient than mechanical stress in contributing to change in cellular volume, our results suggest that the water permeation due to the capillary pressure may be non-negligible over long times or under strong confinement in simple configurations. 
To examine if this is the case in the physiological processes,  currently we are examining the combined effects of osmotic and mechanical stresses on water permeation in the small Peclet number limit, which is relevant to cellular migration in the presence of both confinement and a solute gradient~\cite{StrokaJiangChenEtAl2014_Cell} and has been studied numerically by Jaeger {\em et al.}~\cite{jae-car-med-try1999}.

### 2021-03-18 vesicle_code/docs/permeableVes/semipermeable_vesicle.tex

The value of $\beta$ can be estimated by first
converting the apparent permeability of a 
membrane~\cite{OlbrichRawiczNeedhamEtAl2000_BJ} to the hydraulic
conductivity $k_w$ using the conversion
in~\citet{FettiplaceHaydon1980_PhysRev}, and then scaling $k_w$ to $R_0/\mu$. We found that for most lipid bilayer membranes, $10^{-14}\le
k_w\le 10^{-12}$ m$^2$s/kg, thus the range of $\beta$ corresponding to the native water permeability of a lipid bilayer membrane (vesicle) is $10^{-11}\le
\beta\le 10^{-9}$, consistent with the value in \citet{vogl2014effect}. 
The cellular nuclear envelop is often treated as a filtration membrane \cite{Shahin2019_AdvSci}, and 
%Treating the cellular nuclear envelop as an ultrafiltration membrane, 
its hydraulic conductivity is reported to be order of magnitude larger \cite{Wijmans1984_JMS}:
$10^{-11}\le k_w \le 10^{-9}$m$^2$s/kg,  and the corresponding range of $\beta$ (using the same $R_0/\mu$ for scaling) is $10^{-8}\le
\beta\le 10^{-6}$.

For the simulations in \S~\ref{sec:free_space} and \S~\ref{sec:confinement} we use a wider range of $\beta$ to elucidate the effects of permeating water flow across a
semipermeable inextensible membrane.
%
For example, we use $10^{-5}\le\beta\le10^{-2}$ in our simulations of a
semipermeable inextensible membrane in free space under various flowing
conditions. We found that the equilibrium shape of a semipermeable
membrane in free space to be independent of $\beta$. Within this range
of $\beta$ we find that $\beta$ only affects the time it takes for the
membrane to reach the equilibrium shape, which is determined by the
background flow (such as a quiescent flow, a linear shear flow and its
shear rate, and a Poiseuille flow and its maximum velocity). From these
results and our analysis of area dynamics, we suspect similar
hydrodynamics of a semipermeable vesicle (with $10^{-11}\le\beta\le
10^{-9}$). 
%
When we use $\beta\ge 0.1$, we find that the semipermeable membrane swells to maximize its water content, independent of the initial reduced 
area and the background flow.
%
%In the simulations we focus on intermediate value of $10^{-5}\le \beta\le 10^{-3}$ to investigate the hydrodynamics of a semipermeable inextensible membrane. 
%

@@

We next compute the area dynamics of a semipermeable vesicle. Since the
area of the vesicle is
\begin{align}
  A(t) = \frac{1}{2}\int_{\gamma} (\xx \cdot \nn)\, ds,
  \label{eqn:area}
\end{align}
its derivative includes terms due to the time derivative of $\xx \cdot
\nn$ and a term due to interface stretching~\cite{lai-tse-hua2008}, but
this latter term is zero because of the local inextensibility condition.
Therefore,
\begin{align}
  \dot{A}(t) =
  \frac{1}{2} \int_{\gamma} (\dot{\xx} \cdot \nn)\, ds  + 
  \frac{1}{2} \int_{\gamma} (\xx \cdot \dot{\nn})\, ds.
\end{align}
Applying integration by parts, the two integrals are identical and we combine them to obtain
\begin{align}
  \dot{A}(t) = \int_\gamma \left(\beta (\ff \cdot \nn)\nn 
    + \SS[\ff]\right) \cdot \nn\, ds 
  = \beta \int_\gamma (\ff \cdot \nn)\, ds,
\end{align}
where we have used the incompressibility of the single-layer potential.
Substituting in $\ff = \ff_\mathrm{ben} + \ff_\mathrm{ten}$ and
repeatedly applying integration by parts, we obtain
\begin{align}
  \dot{A}(t) & = \beta \int_\gamma \left(
    \frac{\kappa^3}{2} - \kappa \Lambda \right) \, ds.
  \label{eqn:areaROC}
\end{align}
%\begin{align}
%  \dot{A}(t) &= \beta \int_\gamma \left( -\xx_{ssss} + 
%    (\sigma \xx_s)_s \right) \cdot \nn \, ds \\
%  &= -\beta \int_\gamma \left(-\xx_{sss} + \sigma \xx_s 
%    \right) \cdot \kappa \xx_s \, ds \\
%  &= -\beta \int_\gamma \left(\xx_{ss} \cdot 
%    (\kappa \xx_s)_s + \kappa \sigma \right) \, ds \\
%  &= -\beta \int_\gamma \left(\kappa (\xx_{ss} \cdot \xx_{ss}) + 
%    \kappa_s (\xx_{ss} \cdot \xx_s) + \kappa \sigma \right) 
%    \, ds \\
%  &  = -\beta \int_\gamma \left(\kappa^3 + \kappa \sigma \right) 
%    \, ds = \beta \int_\gamma \left(
%    \frac{\kappa^3}{2} - \kappa \Lambda \right) \, ds.
%  \label{eqn:areaROC}
%\end{align}
Note that the integrand in Equation~\eqref{eqn:areaROC} is zero when the membrane tension balances the elastic force
\begin{align}
\label{eqn:areaROC2}
\Lambda &= \frac{\kappa^2}{2}=\frac{\kappa_0^2}{2},
%  \;\mathrm{ at }\; \mathrm{equilibrium},
\end{align}
which is consistent with the equilibrium result in
Equation~(\ref{eqn:curveODE}).

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Quiescent Flow} 
\begin{figure*}[htp]
  \centering
  \ifTikz
  \input{figures/relaxationComposite.tikz}
  \else
  \includegraphics{figures/relaxationComposite.pdf}
  \fi
  \caption{\label{fig:relaxationComposite} (a) Snapshots of four
  semipermeable inextensible membranes with $\beta = 10^{-8}$ submerged
  in a fluid with no background flow. Time goes from left to right. The
  red regions correspond to influx and the blue regions correspond to
  efflux. In all cases, the vesicle reaches a steady state circular
  shape. (b) The reduced area of each of the simulations. The particular
  snapshots in (a) occur at the marks along the curve. Three dimensioned
  times are denoted by the dashed lines.}
\end{figure*}
In the absence of a background flow, a semipermeable inextensible
membrane in free space is quantified in terms of its area dynamics and
transition to the equilibrium shape, which is a circular membrane with a
uniform membrane tension $\Lambda = 2\pi^2/L^2$. This is in stark
contrast with the impermeable vesicle relaxing to equilibrium that
depends on the vesicle reduced area, which remains constant for
$\beta=0$. Our analysis in \S~\ref{subsec:area_dynamics} shows that a
semipermeable inextensible membrane relaxes to a circle at a rate that
is proportional to $\beta$, regardless of the initial membrane shape
(for a fixed length). For this reason, our numerical results are
reported in terms of the scaled time $\beta t$.

>>>>

The value of $\beta$ can be estimated by first
converting the apparent permeability of a 
membrane~\cite{OlbrichRawiczNeedhamEtAl2000_BJ} to the hydraulic
conductivity $k_w$ using the conversion
in~\citet{FettiplaceHaydon1980_PhysRev}, and then scaling $k_w$ to $R_0/\mu$. We found that for most lipid bilayer membranes, $10^{-14}\le
k_w\le 10^{-12}$ m$^2$s/kg, thus the range of $\beta$ corresponding to the native water permeability of a lipid bilayer membrane (vesicle) is $10^{-11}\le
\beta\le 10^{-9}$, consistent with the value in \citet{vogl2014effect}. 
The cellular nuclear envelop is often treated as a filtration membrane \cite{Shahin2019_AdvSci}, and 
%Treating the cellular nuclear envelop as an ultrafiltration membrane, 
its hydraulic conductivity is reported to be orders of magnitude larger~\cite{Wijmans1984_JMS}:
$10^{-11}\le k_w \le 10^{-9}$m$^2$s/kg,  and the corresponding range of $\beta$ (using the same $R_0/\mu$ for scaling) is $10^{-8}\le
\beta\le 10^{-6}$.

For the simulations in \S~\ref{sec:free_space} and \S~\ref{sec:confinement} we use a wider range of $\beta$ to elucidate the effects of permeating water flow across a
semipermeable inextensible membrane.
%
For example, we use $10^{-5}\le\beta\le10^{-2}$ in our simulations of a
semipermeable inextensible membrane in free space under various flowing
conditions. We found that the equilibrium shape of a semipermeable
membrane in free space to be independent of $\beta$. Within this range
of $\beta$ we find that $\beta$ only affects the time it takes for the
membrane to reach the equilibrium shape, which is determined by the
background flow (such as a quiescent flow, a linear shear flow and its
shear rate, and a Poiseuille flow and its maximum velocity). From these
results and our analysis of areal dynamics, we suspect similar
hydrodynamics of a semipermeable vesicle (with $10^{-11}\le\beta\le
10^{-9}$). 
%
When we use $\beta\ge 0.1$, we find that the semipermeable membrane swells to maximize its water content, independent of the initial reduced 
area and the background flow.
%
%In the simulations we focus on intermediate value of $10^{-5}\le \beta\le 10^{-3}$ to investigate the hydrodynamics of a semipermeable inextensible membrane. 
%

@@

We next compute the areal dynamics of a semipermeable vesicle. Since the
area of the vesicle is
\begin{align}
  A(t) = \frac{1}{2}\int_{\gamma} (\xx \cdot \nn)\, ds,
  \label{eqn:area}
\end{align}
its derivative includes terms due to the time derivative of $\xx \cdot
\nn$ and a term due to interface stretching~\cite{lai-tse-hua2008}, but
this latter term is zero because of the local inextensibility condition.
Therefore,
\begin{align}
  \dot{A}(t) =
  \frac{1}{2} \int_{\gamma} (\dot{\xx} \cdot \nn)\, ds  + 
  \frac{1}{2} \int_{\gamma} (\xx \cdot \dot{\nn})\, ds.
\end{align}
Applying integration by parts, the two integrals are identical and we combine them to obtain
\begin{align}
  \dot{A}(t) = \int_\gamma \left(\beta (\ff \cdot \nn)\nn 
    + \SS[\ff]\right) \cdot \nn\, ds 
  = \beta \int_\gamma (\ff \cdot \nn)\, ds,
\end{align}
where we have used the incompressibility of the single-layer potential.
Substituting in $\ff = \ff_\mathrm{ben} + \ff_\mathrm{ten}$ and
repeatedly applying integration by parts, we obtain
\begin{align}
  \dot{A}(t) & = \beta \int_\gamma \left(
    \frac{\kappa^3}{2} - \kappa \Lambda \right) \, ds.
  \label{eqn:areaROC}
\end{align}
%\begin{align}
%  \dot{A}(t) &= \beta \int_\gamma \left( -\xx_{ssss} + 
%    (\sigma \xx_s)_s \right) \cdot \nn \, ds \\
%  &= -\beta \int_\gamma \left(-\xx_{sss} + \sigma \xx_s 
%    \right) \cdot \kappa \xx_s \, ds \\
%  &= -\beta \int_\gamma \left(\xx_{ss} \cdot 
%    (\kappa \xx_s)_s + \kappa \sigma \right) \, ds \\
%  &= -\beta \int_\gamma \left(\kappa (\xx_{ss} \cdot \xx_{ss}) + 
%    \kappa_s (\xx_{ss} \cdot \xx_s) + \kappa \sigma \right) 
%    \, ds \\
%  &  = -\beta \int_\gamma \left(\kappa^3 + \kappa \sigma \right) 
%    \, ds = \beta \int_\gamma \left(
%    \frac{\kappa^3}{2} - \kappa \Lambda \right) \, ds.
%  \label{eqn:areaROC}
%\end{align}
Note that the integrand in Equation~\eqref{eqn:areaROC} is zero when the membrane tension balances the elastic force
\begin{align}
\label{eqn:areaROC2}
\Lambda &= \frac{\kappa^2}{2}=\frac{\kappa_0^2}{2},
%  \;\mathrm{ at }\; \mathrm{equilibrium},
\end{align}
which is consistent with the equilibrium result in
Equation~(\ref{eqn:curveODE}).

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Quiescent Flow} 
\begin{figure*}[htp]
  \centering
  \ifTikz
  \input{figures/relaxationComposite.tikz}
  \else
  \includegraphics{figures/relaxationComposite.pdf}
  \fi
  \caption{\label{fig:relaxationComposite} (a) Snapshots of four
  semipermeable inextensible membranes with $\beta = 10^{-8}$ submerged
  in a fluid with no background flow. Time goes from left to right. The
  red regions correspond to influx and the blue regions correspond to
  efflux. In all cases, the vesicle reaches a steady state circular
  shape. (b) The reduced area of each of the simulations. The particular
  snapshots in (a) occur at the marks along the curve. Three dimensioned
  times are denoted by the dashed lines.}
\end{figure*}
In the absence of a background flow, a semipermeable inextensible
membrane in free space is quantified in terms of its areal dynamics and
transition to the equilibrium shape, which is a circular membrane with a
uniform membrane tension $\Lambda = 2\pi^2/L^2$. This is in stark
contrast with the impermeable vesicle relaxing to equilibrium that
depends on the vesicle reduced area, which remains constant for
$\beta=0$. Our analysis in \S~\ref{subsec:area_dynamics} shows that a
semipermeable inextensible membrane relaxes to a circle at a rate that
is proportional to $\beta$, regardless of the initial membrane shape
(for a fixed length). For this reason, our numerical results are
reported in terms of the scaled time $\beta t$.
